Evolutionary rescue by aneuploidy in tumors exposed to anti-cancer drugs

### 2025-01-22 ResponseLetter.tex

1. You do an incredible job of walking the reader through this manuscript, making clear your definitions of important concepts, like evolutionary rescue, along the way.

\textcolor{blue}{Thank you for your kind feedback! We are glad our explanations were clear and helpful.}

Two places where you didn't do that, but that could really push this paper over the top into perfection territory, are the following: 1) offering your reader an explanation, even if brief, of the multi-type branching method central to your analyses/models, and why it was the right selection for this endeavor among other methods or evolutionary modeling frameworks used in the past or available to you,

\textcolor{blue}{ % change 53 
We have added a comment in line 239: ``We assume independence between clonal lineages starting from an initial population of $N$ sensitive cells (we check the effect of density-dependent growth on our results below) and therefore, we use multi-type branching processes to model the dynamics of the cancer cell population. Multi-type branching process models the growth and evolution of populations with distinct types, capturing dynamics like mutation and selection. In evolutionary rescue, it predicts the probability of population survival by tracking adaptive mutations that emerge and spread under environmental stress.''
}

@@

2) putting something real-world up against the parameter ranges you used to start the models, or just a brief statement about how they were established from the literature you drew upon (i.e., from experimental tests, patient data, other models, etc.). The latter is just in an effort to put something real to the numerical ranges that a reader can relate to before launching into the modeling.

\textcolor{blue}{ % change 54
We have updated the ``Parameterization'' section (line 159): ``We parametrize most of our simulations by considering melanoma cells and rely on Rew and Wilson (2000) and Bozic et al. (2013) for the division and death rates, respectively.  
Rew and Wilson (2000)  report \emph{in vivo} measurements of the potential doubling times (the waiting time for the number of cells in the tumor to double, disregarding cell death) for a large set of cancer types. The division rate is obtained as $\lambda=\log{2} / T \approx 0.1$ per day. We select this to be the division rate for sensitive and mutant cells. This is a reasonable assumption given that we assume that the mutation makes the cancer cells immune to the chemotherapeutic drugs. 
Bozic et al. (2013) report the growth rate $\Delta_s$ for sensitive melanoma cancer cells, from patient data, from which they deduce the death rate $0.11 \le \mu_s \le 0.17$. We use  $\mu_s=0.14$ per day. Additionally, they observed the growth rate of cancer cells before treatment to be 0.01, which we use as the growth rate of mutant cells, which are resistant to the drug. Thus, we use $\mu_m=0.1-0.01=0.09$ per day as the death rate for mutant cells.'' \\
We also have another paragraph in the same section about parameterization of the model using evidence from breast cancer.
}

@@

2. Do a search for multi-type branching and make sure you consistently use the hyphen or not.

\textcolor{blue}{%change 31
We have made the change to use the hyphen in all mentions of ``multi-type branching'' such that it is consistent with textbooks such as Athreya $\&$ Ney (1972).
}

3. You identify that at least in certain scenarios aneuploidy plays an important role in the evolutionary rescue of tumor cell populations following chemotherapy, at least with one type of drug, and maybe limited to small primary or secondary tumors. But, we never really know the extent that this scenario is applicable. Is it specific to this particular chemotherapy, small tumors, only certain types of cancer? In conjunction with this, there are places in the manuscript where you say things like ``for most parameter values aneuploidy will either play no role in evolutionary rescue or will be the main driver of adaptation without requiring any mutation (one step).'' Both parts of the quoted sentence stand in opposition to one of the more interesting scenarios found in your modeling, which is that aneuploidy can buy a tumor cell population enough time to accumulate a mutation that leads to evolutionary rescue. Like in point 1, making clear the broader implications of your findings, or the limited scenarios in which they might be applicable, pushes this paper over the top into perfection territory because it roots your reader into something relatable, manageable, or applicable to patient care. If the answer is ``I don't know how applicable this finding is or what the broader implications might be,'' you could then add something to the future research section that talks about how this might be addressed experimentally or theoretically. It's ok to not know for sure, but you shouldn't pass the opportunity to talk about it in limits and future research.

\textcolor{blue}{
Thank you for this remark, you are correct to note that this statement is in opposition to some of our results. We have removed this statement and instead, we now write (line 523): ``The `stationary' scenario, in which the aneuploid growth rate is close to zero in the presence of drug, is particularly interesting. Although this may seem like a small region of parameter space in a general model of evolutionary rescue (Fig. 4A, inset), it could be biologically significant for some tumors under specific drug therapies. 
Our results suggest that identifying aneuploidies that are tolerant or stationary may be worthwhile, as these are also expected to enhance the probability of rescue (Figure 3) and extend the window of opportunity for rescue (Figure 6).''\\
We also added the following sentences to the Discussion.
In line 468, emphasis on new text): ``Our results show that the probability of evolutionary rescue increases with the initial tumor size $N$, the drug-sensitive growth rate $\Delta_s$, the mutation rate $v$, and the aneuploidy rate $u$. \emph{The latter indicates that aneuploidy, even when it only provides tolerance, increases the probability that the tumor will be rescued.}'' \\
And in line 493: ``As an example, we have parameterized our model using estimates for melanoma and triple-negative breast cancer (TNBC) cells under drug therapy from the literature. We find that in these cases we are in the third scenario, in which aneuploidy provides at least partial resistance to the drug. It remains to be seen which tumor type and drug combinations produce tolerant and stationary aneuploidies.''\\
The following sentences were already in the Discussion (line 476): ``In this scenario, aneuploidy provides two advantages. First, it delays the extinction of the population, providing more time for the appearance of the resistance mutation. 
Second, it increases the population size relative to a drug-sensitive population, providing more cells in which mutations can occur. 
Together, this increases the cumulative number of mutants that arise.'' and in line 481: ``We find that aneuploidy can significantly affect evolutionary rescue as it reduces the threshold tumor size by at least an order of magnitude even when aneuploidy only provides tolerance.''
}

4. I would revisit some of your figures to make sure groups/categories are not colored in contradictory ways from one graph to the next (i.e., like ``no aneuploidy'' being one color on one graph and a different color in another graph or different categories (``no aneuploidy'' vs. ``aneuploid'' vs. ``stationary'') using the same color in different figures).

@@

\textcolor{blue}{ %change 56
We have changed Figure 6A to be consistent with the other figures and changed the caption for Figure 6 to be consistent with the new coloring.
}

5. The results on detection time associated with tumor size were not intuitive to me on first read. You might want to clarify or rewrite those sentences to make it more clear. It read to me like something about this finding/scenario changed the baseline detection size required to know the tumor cell population was still present and viable.

\textcolor{blue}{%change 48
We have rewritten the sentence about the mean detection time (line 453): ``We find that for small and intermediate-sized tumors the mean detection time is approximately equal to the mean recurrence time (i.e., $\tau_a^r\approx\tau_a^{M}$ for $N<N_m^*$).''
}

@@

6. Table 1 caption doesn't define what the asterisk means.

\textcolor{blue}{ %change 1
We have added the definition of asterisk in the table.
}

7. You never tell your reader what ODEs is... won't be obvious to everyone. I would just right it out on page 9 where I think is the first time you use it.

\textcolor{blue}{%change 2
We have added an explanation of ODE (line 381): ``In this case the dynamics of the number of mutant cells is deterministic and can therefore be modeled by a system of ordinary differential equations (ODE), which describe how the number of mutant cells changes over time by its time derivative (eq.~D2).''
}

8. Are the tests for standing genetic variation essentially genetic drift? Variation is present and in a tumor cell population wildly fluctuating in size due to treatment you just hit on a ``right'' variant stochastically when tumor size is small (less cells) that is resistant and it's driven to fixation? If so, using genetic drift in this section is rooting language and another way of describing alternative scenarios/hypotheses for tumor rebound.

\textcolor{blue}{In our model, where aneuploidy is deleterious prior to drug treatment, standing variation is generated mis-segragation and purged by selection. Since aneuploids are carry a considerable fitness cost, random drift is not expected to effect standing genetic variation. Standing variation actually increases with the population size, whereas the effect of drift decreases with the population size. We now explain this in line 130: ``We assume that $c$ is not too small (Table 1), so that standing variation is generated by a balance between the generation of anuploids and selection against them, without the effect of genetic drift.''
}

9. Do we know if it's reasonable to assume the division rates between cell types (i.e., ``sensitive'' vs. ``mutant'') are the same? Maybe ground that assumption into something we know, or just say we are not sure if that is reasonable or not. This relates to point 1 and rooting the reader into these parameter values with something real or being honest about the assumption they are equal.

\textcolor{blue}{ % change 40
We now ground this assumption in results from the literature (line 163): ``The division rate is obtained as $\lambda=\log{2} / T \approx 0.1$ per day. We take this to be the division rate for sensitive and assume that mutant cells, which are resistant to therapy, have the same division rate. Indeed, doubling times for tumors after relapse is lower than that of primary tumors for a variety of tissues \citep{tezuka2007growth,rodgers2024glioblastoma} and metastatic melanoma has been shown to grow faster then primary melonoma \citep{carlson2003tumor}.''
}

10. Mid page 7, ``decreases'' is mis-spelled

@@

\textcolor{blue}{%change 27
Fixed. 
}

and I would swap out ``very well'' with something more descriptive where you use it.

\textcolor{blue}{
We changed ``the approximations perform very well'' to ``the approximations are accurate'' in two places.
} 
\\
\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Reviewer $\#$2}

@@

\textcolor{blue}{
We owe a serious debt of gratitude to the reviewer for this comment. They have basically done all the work for us. We have incorporated a review of the literature in the Introduction (lines 55-87), and refer to these papers throughout the Results at the equations mentioned. The only slight changes we have made are omitting the follow-ups to Antia et al 2003 and instead including Iwasa et al 2003 Proc B, which we really should have mentioned before and is arguably the closest thing in the literature to this manuscript.
Because the results mentioned by the reviewer are scattered across multiple papers, and because they all follow from simple branching
process approximations, we re-derive all results here for our model.
}

2. I have two general modelling-choice questions whose answers may substantially influence the results. 1) How heritable is a particular aneuploidy?  2) Are aneuploids with a resistance mutation really equivalent to diploids with a resistance mutation? Can the modeling choices made be empirically justified, or can the model (or results) be generalized to include these complications?

\textcolor{blue}{% change y1
Aneuploidy is indeed heritable. The extra-chromosome is lost at a rate similar to it's gain. A recent paper estimated the per generation loss rate in yeast at 8e-5-6e-4, depending on the specific chromosome (Hose et al. Genetics 2024). However, this transition from aneuploidy to euploidy is rare enough that we can neglect it, because the euploid has lower fitness than the aneuploid in the presence of drug.
As for the fitness difference between euploid and aneuploid mutants, this is a good question and the answer probably varies depending on the drug, the chromosome, and other factors. We now address both your questions in the Discussion, in line 542: ``Furthermore, loss of extra-chromosome may occur at a high rate (for estimates in yeast, see Hose et al. (2024)) and the fitness of a euploid cell with a resistance mutation may be higher than that of an aneuploid cell with the same mutation. Future work could test when these effects have significant effect on the results of our model.''
}

@@

- line numbers would have been helpful

\textcolor{blue}{We have added line numbers.}

- abstract: first sentence could give more precise definition of evolutionary rescue

\textcolor{blue}{%change 32
We changed the first sentence in the abstract: ``Evolutionary rescue occurs when a population, facing a sudden environmental change that would otherwise lead to extinction, adapts through beneficial mutations, allowing it to recover and persist.''
}

@@

\textcolor{blue}{%change 45
We have added reference to the recommended synthetic reviews in line 61.
}

- intro: 'there are three ways for a population to survive ...' - I think this statement would be better placed at the beginning of the paragraph, after which you focus on genetic adaptation

\textcolor{blue}{%change 28
We have  moved the statement at the beginning of the paragraph, line 60. 
}

- intro: `most models focus on the probability that at most one mutation rescues the population' - I disagree with this statement. I think the focus is rather on the probability that at least one mutation rescues the population. That is the logic behind the $1-\exp(-Np)$ form: $\exp(-Np)$ is the probability that no mutations rescue the population. In addition to Wilson et al 2017, Osmond $\&$ Coop 2020 (Genetics) also calculate the probability rescue is soft and the expected number of rescuing mutations in a variety of rescue scenarios.

\textcolor{blue}{%change 41
We have removed the statement ``most models focus...''. Instead, we include a review of the evolutionary rescue literature (lines 55-87).
}

@@

- methods: 'The aneuploid division rate is selected...' - I found the grammar odd here.

\textcolor{blue}{%change 36
We repharased to (line line 175): ``For the aneuploid growth rate to be intermediate between those of sensitive and mutant cells, $\Delta_s\ll\Delta_a\ll\Delta_m$, we set the aneuploid division rate to be $0.06 \le \lambda_a \le 0.1$.''
}

- methods: `We estimate c=...' - It sounds like you will estimate c from the observed fraction of aneuploid cells but that is not what you describe below.

\textcolor{blue}{%change 5
We revised the text to make it clear that we estimate $c$ from an experimental result and $f$ from $c$ (line 208): ``To estimate fitness cost of aneuploidy, $c$, we note that Lukow et al. (2021) mixed sensitive and aneuploid A375 melanoma cells at $1:1$ ratio, cultured them in a drug-free environment, and observed the ratio evolve as a function of time with the aneuploid cells declining to $15\%$ after 24 days. Thus, our estimate for the fitness cost is $c=\abs{\log\left(0.15/(1-0.15)\right)/24}\approx0.07$ per day (Chevin 2011).
We estimate the fraction of aneuploid cancer cells in the pre-treatment environment using the formula $f=\tilde{u}\lambda_s / c$, which produces an estimate of $f=10^{-3}\times10^{-1}/0.07$, that is, $0.14\%$ of pre-treatment cancer cells have the beneficial aneuploidy of interest.''
}

@@

- results: I thought it was interesting that the second scenario in eqn 5 was the geometric mean of the other two scenarios. Is there some intution for why this is so?

\textcolor{blue}{We agree, but we couldn't come up with such intuition.}

- results: I thought it might be helpful to give a little more intution about the effect of simultaneously increasing the aneuploid birth and death rates (paragraph starting below eqn 5). The reason why this leads to a decrease in $T^*$ and pushes the system towards the second scenario is that the mutation rate has increased, meaning a lineage doesn't have to persist for as long to almost surely lead to a successful rescue mutant, while the expected persistence time hasn't changed, making it more likely that a rare long-lived aneuploid lineage will lead to rescue.

\textcolor{blue}{%change 51
We have added the recommended comment (line 303): ``Interestingly, increasing both the aneuploid division rate, $\lambda_a$, and the aneuploid death rate, $\mu_a$, such that the growth rate $\Delta_a$ remains constant, leads to a decrease in $T^*$, pushing the system to the second scenario. This is because increasing $\lambda_a$ causes a decrease in $T^*$ as  it increases the effective mutation rate $v\lambda_a$ (as mutations mostly occur during division) and a lineage does not have to survive as long in order to generate a successful mutant.''
}

@@

- results: I don't get the logic at the end of this paragraph, that your results are consistent with aneuploidy conferring resistance by decreasing the division rate. My understanding is that your result says that, holding aneuploid growth rate constant, increasing aneuploid division rate helps until you reach scenario 2 (effectively stationary/critical aneuploids) but beyond that doesn't help. I don't see how reducing division rate helps rescue.

\textcolor{blue}{%change 52
We have removed the sentence about these results being consistent with experimental finding.
}

- results: if $N_a^*$ is similar to the detection threshold, then you predict that any tumour detected is very likely (eg, $>50\%$) to evolve resistance? Does this fit with data on resistance rates?

\textcolor{blue}{ % change 58
We believe that our model does fit with data on resistance rates, and the text now reads (line 316): ``Interestingly, the threshold between small and intermediate tumors, $N_a^*$, is similar to the tumor detection threshold of $4.19 \times 10^6$ cells for a wide variety of tumors~\citep{avanzini2019cancer}. We note that vemurafenib-treated melanomas (i.e. melanomas with sizes above the detection threshold) have a probability $>50\%$ to relapse \citep{piejko2023long,handa2022long}.''
}

@@

- results: much of the discussion of the results (eg below eqn 7) is about the probability of rescue, but the equations all give the inverse of this, critical tumour sizes. I find it a little annoying to keep converting between them, and presenting the probabilities of rescue would also help compare with previous papers. Either way is fine but perhaps something to think about.

\textcolor{blue}{
We agree that having both quantities can be confusing. But we think that in the cotext of cancer evolutionary rescue, the tumor size has a more intuitive interpretation than the rescue probability.
}

- results: I think the rescue times given in the main text (eqn 8) could be explained a little more. For small tumours, this is the 2-step version of the 1-step rescue time given in Orr $\&$ Unckless 2014 (the expectation of their eqn 18), $-1/r$. I.e., the faster the population declines the earlier rescue occurs (conditioned on it occurring), because that is when most mutations happen. For larger tumours this is just a deterministic constant population size result, where the successful mutation is produced at a constant rate.

\textcolor{blue}{%change 46
We have added the following text in line 393: ``For small tumors ($ N \ll N_a^*$), the mean rescue time is the two-step equivalent of the one-step result from Orr and Unckleass (2014; expectation of eq.~18). The mean rescue time (conditioned on rescue occuring) is a decreasing function of the sensitive and aneuploid growth rates and independent of the other model parameters, including tumor size (blue line in Figure S6). This is because if the population rapidly declines but is then rescued, than the resistance mutation must have appeared early; if the population slowly declines, than mutations can appear later and the mean time will be longer. In our focus parameter regime, we have $\Delta_s=-0.04$ and $\Delta_a=-10^{-4}$, such that the mean rescue time is mainly determined by the aneuploid growth rate, $\tau_a \approx 10^4$ days. 
\\
For large tumors ($N \gg N_m^*$), the dynamics are equivalent to a scenario where rescue mutations appear at a constant rate, and the mean rescue time is independent of the aneuploid parameters ($u$, $\lambda_a$, and $\Delta_a$). Increasing the per division mutation rate, $v$, leads to the faster appearance of a rescue mutations and hence reduced mean rescue time. Finally, increasing the tumor size leads to shorter mean rescue time, as more sensitive cells can mutate to become resistant.''
}

- results: `aneuploidy promotes evolutionary rescue after $1/\Delta_a$' $\rightarrow$ `aneuploidy promotes evolutionary rescue after $1/\Delta_s$'

@@

\textcolor{blue}{Fixed.} % change 37

- results: in eqn 9 the variable $p_m$ is used, but it is not defined in the main text

### 2025-01-22 ms.tex
reformatting appendix D

This helps us distinguish between two scenarios for the proliferation time. Firstly, when we have at most one lineages which rescues the cancer cell population:
\begin{equation*}
N\ll N_a^*.
\end{equation*}
As a result, the recurrence time is given by~\citep{avanzini2019cancer}:
\begin{align}\label{meanproliferationtime}
\tau_a^r&\approx\tau_a+\frac{\log p_mN}{\Delta_m}.
\end{align}
The factor of $p_m$ in the second term of \cref{meanproliferationtime} is due to the fact that the lineage is conditioned to survive genetic drift and the time to reach $N$ is shorter then the scenario without this property \citep{orr2014population,smith1974hitch}. % change 35

The second scenario is when the sensitive population produces a large number of rescue lineages in a short period of time. This is given by the condition:
\begin{align*}
N\gg N_a^*.
\end{align*}
As a result, the recurrence time is obtained by solving the following system of ODEs:
\begin{equation}\label{detODE}
\begin{aligned}
\frac{ds}{dt}&=\Delta_ss,\\
\frac{da}{dt}&=\Delta_aa+u\lambda_ss,\\
\frac{dm}{dt}&=\Delta_mm+v\lambda_aa+v\lambda_ss.
\end{aligned}
\end{equation}
Solving the system of ODEs for initial condition $\left(s(0), a(0), m(0)\right)=\left(N,0,0\right)$ we obtain:
\begin{align*}
m\left(t\right)=\frac{Nuv\lambda_a\lambda_s}{\Delta_s-\Delta_a}\left[\frac{\e^{\Delta_st}-\e^{\Delta_mt}}{\Delta_s-\Delta_m}-\frac{\e^{\Delta_at}-\e^{\Delta_mt}}{\Delta_a-\Delta_m}\right]+Nv\lambda_s\frac{\e^{\Delta_st}-\e^{\Delta_mt}}{\Delta_s-\Delta_m}.
\end{align*}
We obtain $\tau_a^r$ such that $m\left(\tau_a^r\right)=N$ by solving:
\begin{equation}\label{eq:t2det2024}
1=\frac{uv\lambda_a\lambda_s}{\Delta_s-\Delta_a}\left[\frac{\e^{\Delta_s\tau_a^r}-\e^{\Delta_m\tau_a^r}}{\Delta_s-\Delta_m}-\frac{\e^{\Delta_a\tau_a^r}-\e^{\Delta_m\tau_a^r}}{\Delta_a-\Delta_m}\right]+v\lambda_s\frac{\e^{\Delta_s\tau_a^r}-\e^{\Delta_m\tau_a^r}}{\Delta_s-\Delta_m}.
\end{equation}
This is a transcendental equation which cannot be solved exactly but we can obtain an approximation by noting that for large $\tau_a^r$ the above equation can be written as: 
\begin{equation*}
1=v\lambda_s\frac{\e^{\Delta_m\tau_a^r}}{\abs{\Delta_s-\Delta_m}},
\end{equation*}
which has solution:
\begin{equation}\label{eq:t2det}
\tau_a^r\approx\frac{1}{\Delta_m}\log\frac{\Delta_m-\Delta_s}{v\lambda_s}.
\end{equation}
We observe that the terms given by the evolutionary trajectory $sensitive \rightarrow aneuploid \rightarrow mutant$ do not contribute to the above approximation and, as a result, we deduce that it is accurate only for $N\gg N_m^*>N_a^*$. %change 9

>>>>

This helps us distinguish between two scenarios for the proliferation time. Firstly, when we have at most one lineage which rescues the cancer cell population:
\begin{equation*}
N\ll N_a^*.
\end{equation*}
As a result, the recurrence time is given by~\citep{avanzini2019cancer}:
\begin{align}\label{meanproliferationtime}
\tau_a^r&\approx\tau_a+\frac{\log p_mN}{\Delta_m}.
\end{align}
The factor of $p_m$ in the second term of \cref{meanproliferationtime} is due to the fact that the lineage is conditioned to survive genetic drift and the time to reach $N$ is shorter then the scenario without this property \citep{orr2014population,smith1974hitch}. % change 35

The second scenario is when the sensitive population produces a large number of rescue lineages in a short period of time. This is given by the condition:
\begin{align*}
N\gg N_a^*.
\end{align*}
As a result, the recurrence time is obtained by solving the following system of ODEs:
\begin{equation}\label{detODE}
\begin{aligned}
\frac{ds}{dt}&=\Delta_ss,\\
\frac{da}{dt}&=\Delta_aa+u\lambda_ss,\\
\frac{dm}{dt}&=\Delta_mm+v\lambda_aa+v\lambda_ss.
\end{aligned}
\end{equation}
Solving the system of ODEs for initial condition $\left(s(0), a(0), m(0)\right)=\left(N,0,0\right)$ we obtain:
\begin{equation} \label{m_t}
m(t) = \frac{N u v \lambda_a \lambda_s}{\Delta_a - \Delta_s}\left[ \frac{\e^{\Delta_m t} - \e^{\Delta_a t}}{\Delta_m - \Delta_a} - \frac{\e^{\Delta_m t} - \e^{\Delta_s t}}{\Delta_m - \Delta_s} \right] + N v \lambda_s \frac{\e^{\Delta_m t} - \e^{\Delta_s t}}{\Delta_m - \Delta_s}.
\end{equation}
We obtain $\tau_a^r$ by solving \eqref{m_t} for the time at which the number of mutant cells reaches $N$, i.e., $m\left(\tau_a^r\right)=N$:
\begin{equation}\label{eq:t2det2024}
1 = \frac{u v \lambda_a \lambda_s}{\Delta_a - \Delta_s}\left[ \frac{\e^{\Delta_m \tau_a^r} - \e^{\Delta_a \tau_a^r}}{\Delta_m - \Delta_a} - \frac{\e^{\Delta_m \tau_a^r} - \e^{\Delta_s \tau_a^r}}{\Delta_m - \Delta_s} \right] + v \lambda_s \frac{\e^{\Delta_m \tau_a^r} - \e^{\Delta_s \tau_a^r}}{\Delta_m - \Delta_s}.
\end{equation}
This is a transcendental equation which cannot be solved exactly but we can obtain an approximation by noting that for large $\tau_a^r$ the above equation can be written as: 
\begin{equation*}
1=v\lambda_s\frac{\e^{\Delta_m\tau_a^r}}{\Delta_m - \Delta_s},
\end{equation*}
which has solution:
\begin{equation}\label{eq:t2det}
\tau_a^r\approx\frac{1}{\Delta_m}\log\frac{\Delta_m-\Delta_s}{v\lambda_s}.
\end{equation}
We observe that the terms given by the evolutionary trajectory \textit{sensitive} $\rightarrow$ \textit{aneuploid} $\rightarrow$ \textit{mutant} do not contribute to the above approximation and, as a result, we deduce that it is accurate only for $N \gg N_m^* > N_a^*$. %change 9
